---
title: "Bootstrapping Convergence"
author: "Aris Paschalidis"
format: html
editor: visual
execute:
  cache: true
---

```{r install pkgs}
#| include: FALSE
library(coiaf)
library(ggplot2)
```

In this document, we explore the convergence of bootstrapping replicates. In other words, we aim to identify how many replicates are needed before our bootstrapping functions produce a relatively small confidence interval. We first run our bootstrapped functions with a varying number of replicates.

```{r bootstrap}
#| warning: FALSE
sim_data <- sim_biallelic(coi = 5, plmaf = runif(1000, 0, 0.5))
n_bootstraps <- c(seq(10, 100, 10), seq(200, 1000, 100))
bootstraps <- purrr::map_dfr(n_bootstraps, function(rep) {
  bootstrap_ci(sim_data, replicates = rep)
}) %>% 
  tibble::add_column(n_bootstraps = n_bootstraps)
```

Following running our algorithm, we can next plot our data with the x-axis representing the number of bootstrap replicates used and the y-axis representing the standard error.

```{r convergence}
bootstraps %>% 
  dplyr::filter(std.error != 0) %>% 
  ggplot(aes(x = n_bootstraps, y = std.error)) +
  geom_point() +
  geom_line() +
  scale_x_continuous(breaks = seq(0, 1000, 100)) +
  labs(y = "Standard Error", x = "Number of Bootstraps") +
  theme_coiaf()
```

Looking at the figure produced, we note that after 100 bootstrap iterations there is little difference in the standard error, even as the number of bootstraps increases. As a result, we can set the default number of replicates in our bootstrapping function to 100.

